Fractional Types

### 2013-02-05 frac.tex
cat background

Even though the categorical semantics of {{langRevTF}} is not the main focus
of this paper, we find it helpful to briefly outline an interpretation of
{{langRevTF}} in \emph{dagger compact closed categories}. These categories
were introduced by Abramsky and
Coecke~\cite{Abramsky:2004:CSQ:1018438.1021878} and then by
Selinger~\cite{Selinger:2007:DCC:1229185.1229207}. The simplest example of
such categories is the category of finite sets and relations on which we
focus.

the category of finite sets and relations.

@@

>>>>

Even though the categorical semantics of {{langRevTF}} is not the main focus
of this paper, we find it helpful to briefly outline an interpretation of
{{langRevTF}} in \emph{biproduct dagger compact closed categories}. These
categories were introduced by Abramsky and
Coecke~\cite{Abramsky:2004:CSQ:1018438.1021878} and then by
Selinger~\cite{Selinger:2007:DCC:1229185.1229207}. The bare definitions are
collected in Appendix~\ref{app:cat}.

The simplest example of such categories is the category of finite sets and
relations on which we focus. To interpret {{langRevTF}} in this category, we
associate a set with each type. For {{1}}, we associate a set with one
element and we interpret {{*}} as the cartesian product and {{+}} as the
disjoint union. For {{1/b}}, we associate the same set associated with {{b}}
and hence, in our denotations, we identify {{b}} and {{1/b}} as well as {{v}}
and {{1/v}}. It is intuitively clear that this category is compact closed as
any relation between $A$ and $B$ can be represented as a finite set of type
{{A * B}}. For example, consider the ``boolean negation'' relation:

{{ {[ (1/left(), right()), (1/right (), left ()) ]} }}

This is clearly isomorphic to the corresponding set of pairs. The map
{{eta_b}} constructs the identity relation at type {{b}}. For example, {{eta_{1+1} }}
is the relation:

{{ {[ (1/(), (left(), left())), (1/(), (right(),right ())) ]} }}

@@

\appendix
\section{Categorical Background}
\label{app:cat}

A \emph{dagger category} is a category together with an involutive,
identity-on-objects, contravariant functor $\dagger$. Concretely, this means
that to every morphism $F : A \rightarrow B$, one associates a morphism
$f^{\dagger} : B \rightarrow A$, called the \emph{adjoint} of $f$, such that
for all $f : A \rightarrow B$ and $g : B \rightarrow C$, we have:
\[\begin{array}{rcl}
\mathit{id}^\dagger_A &=& \mathit{id}_A \\
(g \circ f)^\dagger &=& f^\dagger \circ g^\dagger \\
(f^dagger)^\dagger &=& f
\end{array}\]

A \emph{dagger symmetric monoidal category} is a symmetric monoidal category
with a dagger structure such that the contravariant functor $\dagger$
coherently preserves the symmetric monoidal structure. Concretely, this
requirement means that for all $f : A \rightarrow B$ and $g : C \rightarrow
D$, we have:
\[\begin{array}{rcl}
(f \otimes g)^\dagger &=& f^\dagger \otimes g^\dagger \\
\alpha^\dagger_{A,B,C} &=& \alpha^{-1}_{A,B,C} \\
\lambda^\dagger_A &=& \lambda^{-1}_A \\
\sigma^\dagger_{A,B} &=& \sigma^{-1}{A,B}
\end{array}\]

\begin{definition}[Dagger Compact Closed Category]
\label{def:cat}
A \emph{dagger compact closed category} is a dagger symmetric monoidal
category that is also compact closed and such that for all $A$:
\[
\eta_A = \sigma_{A,A^*} \circ \epsilon^\dagger_A
\]
\end{definition}

\todo{biproducts}
